Red Wine Exolloratory Data Analysis by Youwu Mike Lu
========================================================
```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(dplyr)
library(ggplot2)
library(ggplot2)
library(GGally)
library(scales)
library(memisc)
library(MASS)
```

```{r echo=FALSE}
# Load the Data
rwine <- read.csv("wineQualityReds.csv")
```

This report explores a datasert containing different chemical properties as well
as the quality of the wines.

This report tries to understand if the chemical properties of the red wines
play a role in its quality.

```{r echo=FALSE}
#Taking a look at the dataset
head(rwine)
```

```{r}
names(rwine)
```


```{r echo=FALSE, Univariate_Plots}
dim(rwine)
```

```{r}
str(rwine)
```

```{r}
#Another quick summarized look at the dataset
summary(rwine)
```

There are 1,599 observations with 13 columns. The first column, X, does not
contain any variable but simple the numbering. Therefore, there are only 12
variables in the datset. 

Quality will be the main focus of this analysis. 

```{r echo=FALSE}
rwine$X <- NULL
```

In order to proceed, first by removing variable X.

```{r echo=FALSE}
head(rwine)
```

```{r echo=FALSE}
summary(rwine$quality)
```

To better categorize the quality if the wine, let's separate them into 3
categories: undesirable, good and amazing.

In order to do so, let's first transform quality from integers to an ordered
factor. 

```{r}
# Transform quality from an integer to an ordered factor
rwine$quality <- factor(rwine$quality, ordered = T)
```

```{r}
#Create the rating variable
rwine$rating <- ifelse(rwine$quality < 5, 'undesirable', ifelse(
  rwine$quality < 7, 'good', 'amazing'))
```

```{r}
#Ordering the new variable rating
rwine$rating <- ordered(rwine$rating,
                     levels = c('undesirable', 'good', 'amazing'))
```

```{r echo=FALSE}
#Create plot for quality
qplot(data = rwine, x = quality) + 
  ggtitle('Quality')
```

As seen from the graph, and confirmed from the data analysis, most of the wine
quality are between 5 and 6, and none of them are either 1,2 or 9.

```{r echo=FALSE}
#Create plot for rating
qplot(data = rwine, x = rating) + 
  ggtitle('Rating')
```


When plotting for rating, it is clear that once again, most wines fall within
the good category, theer are a few undesirable wines with a bit more of amazing
wines. 

When talking about wine, one of the first thing to analyze at is the alcohol 
level!

```{r}
#Get summary for variable alcohol
summary(rwine$alcohol)
```

```{r echo=FALSE}
#Create plot for alcohol
qplot(data = rwine, x = alcohol, binwidth = 0.2) + 
  ggtitle('Alcohol Content')
```

```{r echo=FALSE}
#Creating boxplot for alcohol
ggplot(data=rwine, aes(y=alcohol)) +
  geom_boxplot() + ggtitle('Alcohol Content Box Plot')
```

The alcohol content has a right skewed distribution, where most of them tends
to be lower. The average is 10.42. There are a few outliers, with the highest
being 14.90.

Next, let's explore the other variables going down their correlation level.

```{r}
summary(rwine$fixed.acidity)
```

```{r echo=FALSE}
ggplot(data=rwine, aes(y=fixed.acidity)) +
  geom_boxplot() + ggtitle('Fixed Acidity Box Plot')
```

```{r echo=FALSE}
qplot(data = rwine, x = fixed.acidity, binwidth = 0.2) + 
  ggtitle('Fixed Acidity Plot')
```

This is a normally distributed variable, most wines have their fixed acidity
level beween 7 and 9, with a few outliers over 13, all the way up to 15.9.

```{r}
summary(rwine$sulphates)
```

```{r echo=FALSE}
ggplot(data=rwine, aes(y=sulphates)) +
  geom_boxplot() + ggtitle('Sulphates Level Box Plot')
```

```{r echo=FALSE}
qplot(data = rwine, x = sulphates, binwidth = 0.02) + 
  ggtitle('Sulphates Level Plot')
```

The sulphate level is right skewed with mean of 0.65 and some outliers ranging
up to 2.

```{r}
summary(rwine$citric.acid)
```

```{r echo=FALSE}
ggplot(data=rwine, aes(y=citric.acid)) +
  geom_boxplot() + ggtitle('Citric Acid Level Box Plot')
```

```{r echo=FALSE}
qplot(data = rwine, x = citric.acid, binwidth = 0.02) + 
  ggtitle('Citric Acid Level Plot')
```

Most of the citric acid level are under 0.75, to be more specific, 75% are under
0.42. There is an outlier with a value of 1.

```{r}
summary(rwine$total.sulfur.dioxide)
```

```{r echo=FALSE}
ggplot(data=rwine, aes(y=total.sulfur.dioxide)) +
  geom_boxplot() + ggtitle('Total Sulfur Dioxide Level Box Plot')
```

```{r echo=FALSE}
qplot(data = rwine, x = total.sulfur.dioxide, binwidth = 2) + 
  ggtitle('Total Sulfur Dioxide Level Plot')
```

Once again, the total sulfur level is right skewed, with most of them falling
between 22 and 62.

```{r}
summary(rwine$density)
```

```{r echo=FALSE}
ggplot(data=rwine, aes(y=density)) +
  geom_boxplot() + ggtitle('Density Level Box Plot')
```

```{r echo=FALSE}
qplot(data = rwine, x = density, binwidth = 0.0002) + 
  ggtitle('Density Level Plot')
```

The density is normally distributed, with a mean of 0.9968.

```{r}
summary(rwine$chlorides)
```

```{r echo=FALSE}
ggplot(data=rwine, aes(y=chlorides)) +
  geom_boxplot() + ggtitle('Chlorides Level Box Plot')
```

```{r echo=FALSE}
qplot(data = rwine, x = chlorides, binwidth = 0.005) + 
  ggtitle('Chlorides Level Level Plot')
```

The chlorides level is right skewed. Discounting the outliers, we can see that
most of the values are actually normally distributed with a mean of 0.079.
Using the log10 of the distribution to dampen the effect of the ouotliers, we
have the following graph.

```{r echo=FALSE}
ggplot(aes(x=chlorides), data=rwine) +
  coord_cartesian(xlim=c(0.05,0.12)) +
  geom_histogram(binwidth=0.005) +
  scale_x_log10(breaks=seq(0.07,0.12,0.07)) +
  ggtitle("Log10(Chlorides) Level Plot")
```

This graph confirms the finding that by removing the outliers, the chlorides
level are normally distributed around its mean.

```{r}
summary(rwine$fixed.acidity)
```

```{r echo=FALSE}
ggplot(data=rwine, aes(y=fixed.acidity)) +
  geom_boxplot() + ggtitle('Fixed Acidity Level Box Plot')
```

```{r echo=FALSE}
qplot(data = rwine, x = fixed.acidity, binwidth = 0.2) + 
  ggtitle('Fixed Acidity Level Level Plot')
```

The acidity is normally distributed with a mean of 7.90. As seen on the box 
plot, there is still a few outliers.

```{r}
summary(rwine$pH)
```

```{r echo=FALSE}
ggplot(data=rwine, aes(y=pH)) +
  geom_boxplot() + ggtitle('pH Level Box Plot')
```

```{r echo=FALSE}
#Create plot for fixed.acidity
qplot(data = rwine, x = pH, binwidth = 0.02) + 
  ggtitle('pH Level Level Plot')
```

Same with the fixed acidity level, the pH level is normally distributed. Having
a mean of 3.10, it can be seen that red wines are acidic.

```{r}
summary(rwine$residual.sugar)
```

```{r echo=FALSE}
ggplot(data=rwine, aes(x='',y=residual.sugar)) +
  geom_boxplot() + ggtitle('Residual Sugar Level Box Plot')
```

```{r echo=FALSE}
qplot(data = rwine, x = residual.sugar, binwidth = 0.2) + 
  ggtitle('Residual Sugar Level Level Plot')
```

Same with chloride level, on first look, the residual sugar seems to be a right
skewed distribution. But with the help of the box plot, we can see that there 
are many outliers. By removing them, we will have another perspective. 

```{r echo=FALSE}
ggplot(aes(x=residual.sugar), data=rwine) +
  coord_cartesian(xlim=c(0.6,10)) +
  geom_histogram(binwidth=0.02) +
  scale_x_log10(breaks=seq(0,20,2)) +
  ggtitle("Log10(Residual Sugar Level Level Plot)")
  
```

From this new plot, it can be seen that the residual sugar level is indeed 
normally distributed around its mean.

```{r}
summary(rwine$free.sulfur.dioxide)
```

```{r echo=FALSE}
ggplot(data=rwine, aes(y=free.sulfur.dioxide)) +
  geom_boxplot() + ggtitle('Free Sulfur Dioxide Level Box Plot')
```

```{r echo=FALSE}
qplot(data = rwine, x = free.sulfur.dioxide, binwidth = 1) + 
  ggtitle('Free Sulfur Dioxide Level Level Plot')
```

The free sulfur dioxide level is right skewed. Given that the distance
between the first quartile and median is shorter than the distance between the
median the the third quartile,  it can be concluded that the free sulfur dioxide
level tends to concentrate more in Q2.

# Univariate Analysis

### What is the structure of your dataset?

There are 1,599 observations of the red wine data. The quality of these wines 
have been rated by 3 wine experts from 0 (very bad) to 10 (excellent). There is
a total of 12 variables, 11 input variables and 1 output variable.

The input variables are:
- alcohol;
- chlorides;
- citric acid;
- fixed acidity;
- free sulfur dioxide;
- pH;
- residual sugar;
- sulphates;
- total sulfur dioxide;
- volatile acidity.

Output variable:
- quality

### What is/are the main feature(s) of interest in your dataset?

The main feature of interest in the quality of the wine. More specifically,
is there a common factor for in the wines that are perceived as better quality
by the wine expert, and how does the different chemical properties impact the 
perceived quality.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

There are many features that are interesting to look at. As seen from a brief
correlation analysis above, the first one is the alcohol content, volatile 
acidity, sulphates and citric acid all play a bigger role in the perceived
quality of a wine.

### Did you create any new variables from existing variables in the dataset?

I created a new variable called rating, ranging from "undesirable" (lower than
5), "good" (between 5 and 7), and "amazing" (higher than 7) in order to counter
the lack of spread in the quality rating.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

This is a very clean and tidy dataset. Only the "X" column has been deleted
since it was not useful. The quality variable has been changed to an ordered 
factor for visualization, but to perform further analysis, it has been changed
back to its numerical value. 

Some of the unusual distributions are that chlorides, residual sugar and free 
sulfur dioxide levels are all skewed to the right. The citric acid was also
interesting as it has spikes at different levels, possibly suggesting that
it was made on purpose, perhaps due to a lot of wine producers using the same
kind of grapes.

# Bivariate Plots Section

Taking a look at which variable is correlated with the quality of the wines.

```{r echo=FALSE}
rwine2 <- dplyr::select(rwine,-c(rating))
rwine2$quality <- as.numeric(rwine2$quality)
round(cor(rwine2, rwine2$quality),4)
```

As the correlation analysis demonstrates, alcohol content is the highest 
correlated to quality, followed by volatile acidity and sulphates. It is 
interesting to see how pH seems to not have any impact on the quality of the 
wine, taking a closer look at a few of those variables.

It can be seen that alcohol content has the strongest correlation 
with quality. Therefore, taking a closer look at the alcohol content.

```{r echo=FALSE}
#Plot relationship between quality and alcohol
ggplot(aes(y=alcohol, x=quality), data=rwine) + 
  geom_jitter(alpha = 1/10) +
  geom_boxplot(alpha = 1/10, color = 'blue') +
  stat_summary(fun.y = 'mean', geom = 'point', color = 'red') + 
  labs(x = 'Quality', y = 'Alcohol Content (%)', title = 'Quality vs Alcohol')
```

```{r echo=FALSE}
ggplot(aes(y=alcohol, x=rating), data=rwine) + 
  geom_jitter(alpha = 1/10) +
  geom_boxplot(alpha = 1/10, color = 'blue') +
  stat_summary(fun.y = 'mean', geom = 'point', color = 'red') + 
  labs(x = 'Rating', y = 'Alcohol Content (%)', title = 'Rating vs Alcohol')
```

As seen from the graphs above, there is a clear correlation that whenever
the alcohol content is higher, the quality also tends to be higher. It is even
more evident when plotting against the rating, as this elimaates some of the 
lack of spreads.

```{r echo=FALSE}
ggplot(aes(y=volatile.acidity, x=quality), data=rwine) + 
  geom_jitter(alpha = 1/10) +
  geom_boxplot(alpha = 1/10, color = 'blue') +
  stat_summary(fun.y = 'mean', geom = 'point', color = 'red') + 
  labs(x = 'Quality', y = 'Volatile Acidity', 
       title = 'Quality vs Volatile Acidity')
```

```{r echo=FALSE}
ggplot(aes(y=volatile.acidity, x=rating), data=rwine) + 
  geom_jitter(alpha = 1/10) +
  geom_boxplot(alpha = 1/10, color = 'blue') +
  stat_summary(fun.y = 'mean', geom = 'point', color = 'red') + 
  labs(x = 'Rating', y = 'Volatile Acidity', 
       title = 'Rating vs Volatile Acidity')
```

From the 2 box plots avove, it can be seen that the volatile acidity and 
the quality are inversely correlated. Meaning that a higher volatile acidity
tends to lead to a lower perceived quality. Once again, this finding can be 
seen even more effectively when plotting against the rating.

```{r echo=FALSE}
ggplot(aes(y=citric.acid, x=quality), data=rwine) + 
  geom_jitter(alpha = 1/10) +
  geom_boxplot(alpha = 1/10, color = 'blue') +
  stat_summary(fun.y = 'mean', geom = 'point', color = 'red') + 
  labs(x = 'Quality', y = 'Citric Acid', title = 'Quality vs Citric Acid')
```

```{r echo=FALSE}
ggplot(aes(y=citric.acid, x=rating), data=rwine) + 
  geom_jitter(alpha = 1/10) +
  geom_boxplot(alpha = 1/10, color = 'blue') +
  stat_summary(fun.y = 'mean', geom = 'point', color = 'red') + 
  labs(x = 'Rating', y = 'Citric Acid', title = 'Rating vs Citric Acid')
```

On the quality plot, it can be seen that the citric acid level has a positive
correlation with the quality, where higher level of citric acid tends to lead
to a higher perceived quality level. This can be further reinforced when 
plotting this varriable against the rating variable.

```{r echo=FALSE}
ggplot(aes(y=sulphates, x=quality), data=rwine) + 
  geom_jitter(alpha = 1/10) +
  geom_boxplot(alpha = 1/10, color = 'blue') +
  stat_summary(fun.y = 'mean', geom = 'point', color = 'red') + 
  labs(x = 'Quality', y = 'Sulphates', title = 'Quality vs Sulphates')
```

```{r echo=FALSE}
ggplot(aes(y=sulphates, x=rating), data=rwine) + 
  geom_jitter(alpha = 1/10) +
  geom_boxplot(alpha = 1/10, color = 'blue') +
  stat_summary(fun.y = 'mean', geom = 'point', color = 'red') + 
  labs(x = 'Rating', y = 'Sulphates', title = 'Rating vs Sulphates')
```

The sulphates level is positively correlated with the quality of the wine. 
Interestingly enough, the median as well as mean for quality 7 to 8 are quite 
similar. This can be intepreted as a plateau  that the sulphates level should
not exceed a certain point.

In order to reinforce this assumption, doing another analysis.

```{r echo=FALSE, message=FALSE, , warning=FALSE}
ggplot(aes(x=sulphates, y=quality), data=rwine) + 
  geom_point(position = 'jitter', alpha = 1/4) +
  labs(y = 'Quality', x = 'Sulphates', title = 'Sulphates vs Quality')
```

This above plot reinforces the previous assumption. Once the sulphates level is 
too high, it does not bring any additional positive value to the quality. But 
rather, it lower the perceived quality of a wine.

```{r echo=FALSE}
ggplot(aes(y=pH, x=quality), data=rwine) + 
  geom_jitter(alpha = 1/10) +
  geom_boxplot(alpha = 1/10, color = 'blue') +
  stat_summary(fun.y = 'mean', geom = 'point', color = 'red') + 
  labs(x = 'Quality', y = 'pH', title = 'Quality vs pH')
```

```{r echo=FALSE}
ggplot(aes(y=pH, x=rating), data=rwine) + 
  geom_jitter(alpha = 1/10) +
  geom_boxplot(alpha = 1/10, color = 'blue') +
  stat_summary(fun.y = 'mean', geom = 'point', color = 'red') + 
  labs(x = 'Rating', y = 'pH', title = 'Rating vs pH')
```

According to the correlation analysis, the correlation between quality and pH
tends to be low. From the quality graph, it can be seen that thre is a only a 
slight correlation between pH and quality. But on the rating graph, this 
correlation is more pronounced. It can be noted that the better the quality of a
wine, the lower its pH level tends to be.

```{r echo=FALSE}
ggplot(aes(y=residual.sugar, x=quality), data=rwine) + 
  geom_jitter(alpha = 1/10) +
  geom_boxplot(alpha = 1/10, color = 'blue') +
  stat_summary(fun.y = 'mean', geom = 'point', color = 'red') + 
  labs(x = 'Quality', y = 'Residual Sugar', title = 'Quality vs Residual Sugar')
```

```{r echo=FALSE}
ggplot(aes(y=residual.sugar, x=rating), data=rwine) + 
  geom_jitter(alpha = 1/10) +
  geom_boxplot(alpha = 1/10, color = 'blue') +
  stat_summary(fun.y = 'mean', geom = 'point', color = 'red') + 
  labs(x = 'Rating', y = 'Residuak Sugar', title = 'Rating vs Residual Sugar')
```

There is no clear correlation between the  residual sugar level and the quality
of a wine.

A correlation analysis has been run against the quality of the wines previously.
Now let's take a look at the correlations between the other variables.

```{r echo=FALSE}
round(cor(rwine2),4)
```

Before doing further analysis, taking a look at the above information in a
graphical form.

```{r echo=FALSE, message=FALSE, fig.width=10, fig.height=10}
#ggairs to illustrate the relationship between the variables
ggpairs(dplyr::select(rwine,-c(residual.sugar, density, free.sulfur.dioxide,rating)),
        title = "Red Wine Correlation Matrix", 
        mapping = aes(color = quality),
        lower = list(continuous = wrap("points", shape = I('.'))),
        upper = list(combo = wrap("box", outlier.shape = I('.'))))
```

Let's first start by taking a loot between alcohol, the most relevant indicator
of quality, against its highest correlated variables,  density and pH.

```{r echo=FALSE,message=FALSE, warning=FALSE}
ggplot(data = rwine, aes(x = alcohol, y = density)) +
  labs(y = 'Density', x = 'Alcohol', title = 'Alcohol vs Density') + 
  scale_x_continuous() +
  geom_point(color = 'red', alpha = 0.5) + 
  geom_smooth() 
```

This graph shows that there is a negative correlation between the alcohol level
and the density level. Which is to be expected, as alcohol is less dense than
water.

```{r echo=FALSE,message=FALSE, warning=FALSE}
ggplot(data = rwine, aes(x = alcohol, y = pH)) +
  labs(y = 'pH', x = 'Alcohol', title = 'Alcohol vs pH') + 
  scale_x_continuous() +
  geom_point(color = 'red', alpha = 0.5) + 
  geom_smooth() 
```

As for the pH level vis-a-vis the alcohol level, there is a slight positive 
correlation between them. As the above graph shows, the pH will increase with
the alcohol level to a certain point, and the decrease is caused by a single
outlier which contained high level of alcohol.

```{r echo=FALSE,message=FALSE, warning=FALSE}
ggplot(data = rwine, aes(y = fixed.acidity, x = pH)) +
  labs(y = 'Fixed Acidity', x = 'pH', title = 'pH vs Fixed Acidity') + 
  scale_x_continuous() +
  geom_point(color = 'red', alpha = 0.5) + 
  geom_smooth() 
```


```{r echo=FALSE,message=FALSE, warning=FALSE}
ggplot(data = rwine, aes(y = citric.acid, x = pH)) +
  labs(y = 'Citric Acid', x = 'pH', title = 'pH vs Citric Acid') + 
  scale_x_continuous() +
  geom_point(color = 'red', alpha = 0.5) + 
  geom_smooth() 
```

As expected, the higher the level of citric acidity or fixed acidity, the lower 
the pH. Which is a numbering representation of the acidity or basisity level, 
where a lower number indicates it being more acid. Both the fixed acidity and
citric acidity follow are negatively correlated to pH, therefore we can 
hypothesis that they are positively correlated. Taking a closer look.

```{r echo=FALSE,message=FALSE, warning=FALSE}
ggplot(data = rwine, aes(y = citric.acid, x = fixed.acidity)) +
  labs(y = 'Citric Acid', x = 'Fixed Acidity', 
       title = 'Fixed Acidity vs Citric Acid') + 
  scale_x_continuous() +
  geom_point(color = 'red', alpha = 0.5) + 
  geom_smooth() 
```

It can be clearly seen that there is a positive correlation between fixed
acidity and citric acidity. Once again, it is interesting to note how many
wines have a 0 or close to 0 citric acidity.

```{r echo=FALSE,message=FALSE, warning=FALSE}
ggplot(data = rwine, aes(x = volatile.acidity, y = citric.acid)) +
  labs(y = 'Citric Acid', x = 'Volatile Acidity', 
       title = 'Volatile Acidity vs Citric Acid') + 
  scale_x_continuous() +
  geom_point(color = 'red', alpha = 0.5) + 
  geom_smooth() 
```

From this graph, we can see that there is a strong negative correlation between  
volatile acidity and the citric acid level in the wine. It is interesting to
note that there is somekind of plateau for citric acidity, arunod 0.5.

```{r echo=FALSE,message=FALSE, warning=FALSE}
ggplot(data = rwine, aes(x = volatile.acidity, y = fixed.acidity)) +
  labs(y = 'Fixed Acidity', x = 'Volatile Acidity', 
       title = 'Volatile Acidity vs Fixed Acidity') + 
  scale_x_continuous() +
  geom_point(color = 'red', alpha = 0.5) + 
  geom_smooth() 
```

There is a slight negative correlation between volatile acidity and fixed acidity.

```{r echo=FALSE,message=FALSE, warning=FALSE}
ggplot(data = rwine, aes(y = citric.acid, x = density)) +
  labs(y = 'Citric Acidity', x = 'Density', 
       title = 'Density vs Citric Acid') + 
  scale_x_continuous() +
  geom_point(color = 'red', alpha = 0.5) + 
  geom_smooth() 
```

```{r echo=FALSE,message=FALSE, warning=FALSE}
ggplot(data = rwine, aes(y = fixed.acidity, x = density)) +
  labs(y = 'Fixed Acidity', x = 'Density', 
       title = 'Density vs Fixed Acidity') + 
  scale_x_continuous() +
  geom_point(color = 'red', alpha = 0.5) + 
  geom_smooth() 
```

From the two graphs above, we can see that both the citric acid  and  fixed
acidity are positively correlated to density. Which is expected given that
a high citric acid will lead to a high fixed acidity. And that a high level
of acidity will be more dense than one of lower level.

```{r echo=FALSE,message=FALSE, warning=FALSE}
ggplot(data = rwine, aes(y = sulphates, x = chlorides)) +
  labs(y = 'Sulphates', x = 'Chlorides', 
       title = 'Chlorides vs Sulphates') + 
  scale_x_continuous() +
  geom_point(color = 'red', alpha = 0.5) + 
  geom_smooth() 
```

From this graph, we can see that there is a high concentration of both chlorides
and sulphates at lower level, and there seems to be a positive correlation 
between them.

```{r echo=FALSE,message=FALSE, warning=FALSE}
ggplot(data = rwine, aes(y = sulphates, x = chlorides)) +
  labs(y = 'Sulphates', x = 'Chlorides', 
       title = 'Chlorides vs Sulphates') + 
  scale_x_continuous() +
  geom_point(color = 'red', alpha = 0.5) + 
  xlim(0, 0.2) +
  ylim(0.25, 1.2) +
  geom_smooth() 
```

Once zooned ib,  the positive correlation can be seen a  bit more clearly. 
Although there is somewhat of a dip at the 0.125 chlorides level, the general
trend is still that of a positive correlation.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

The main feature of interest is to see if any variable affected the quality of
the wine.
Some of the positive correlation are citric acid, sulphate and of course 
alcohol level.
While some of the negative correlation are volatile acidity and its insuing pH
level.

More observations:
- Alcohol has the highest correlation with the quality of a wine;
- The residual sugar has no correlation with the quality of a wine;
- Alcohol is not highly correlated to any metrics other than density,  which it
is negatively correlated;
- Volatile acidity is inversely correlated to the quality of a wine, while 
citric acid is positively correlated to it.
- The pH level, which is affected by the volatilve acidity, fixed  acidity and
citric acid,  is negatively correlated  to quality, meaning a more sour wine
will tend to be ranked higher;
- Sulphates is positively correlated to quality, although a too high 
concentration of sulphates will then be considered undesirable.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

As expected, fixed acidity and citric acid as well as pH are all related.
While the volatile acidity and  fixed acidity, as well as volatile acidity and
the citric acid are all inversely correlated. The density also goes up with a
higher level of citric acid and fixed acidity.

Another observation is between alcohol and density, which makes sense given that
alcohol molecules are less dense than water molecules.

Another intersting observation is that the sulphates level and chlorides level
are positively related, with the sulphates playing a direct role in the 
perceived  quality.

### What was the strongest relationship you found?

The strongest relationship is the one of pH and fixed acidity. While the 
alcohol has  the strongest relationship with the quality of the wine.

# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}
ggplot(aes(x=alcohol, y=volatile.acidity, color = quality), data=rwine) + 
  labs(x = 'Alcohol', y = 'Volatile Acidity',
       title = 'Alcohol and Volatile Acidity by Quality') +
  scale_color_brewer(palette="RdYlBu", name = "Quality") +
  geom_point()
```

Starting by lotting the two highest correlated with quality, alcohol and 
volatile acidity, it can be seen that there is a clear positive correlation
between those 2 variables and the perceived quality of the wine. One interesting
side note is that too high of volatile acidity will actually lead to a decrease
in perveied quality of the wine.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=alcohol, y=volatile.acidity), data=rwine) + 
  labs(x = 'Alcohol', y = 'Volatile Acidity',
       title = 'Alcohol and Volatile Acidity by Quality') +
  facet_wrap(~rating) + 
  geom_smooth() +
  scale_color_brewer(palette="RdYlBu", name = "Quality") +
  geom_point(aes(color = quality), position = 'jitter')
```

By further breaking them down by rating, it can be more clearly seen that 
the higher volatile acidity will actually lead to a lower quality in wine. But
despite this, a higher degree of both alcohol and volatile acidity will lead to
a higher perceived quality of the wine.

```{r echo=FALSE}
ggplot(aes(x=alcohol, y=citric.acid, color = quality), data=rwine) + 
  labs(x = 'Alcohol', y = 'Citric Acid',
       title = 'Alcohol and Citric Acid by Quality') +
  scale_color_brewer(palette="RdYlGn", name = "Quality") +
  geom_point()
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(y=alcohol, x=citric.acid), data=rwine) + 
  labs(y = 'Alcohol', x = 'Citric Acid',
       title = 'Citric Acid And Alcohol by Rating') +
  scale_color_brewer(palette="Set1", name = "Rating") +
  geom_smooth(aes(color = rating))+
  geom_point(aes(color = rating), position = 'jitter', alpha = 0.5) 
```

In the above graph, it can be seen that both the alcohol and citric acid will
positively impact the quality of a wine. However, alcohol will have a
bigger impact on quality. Citric acid plays an important role, as most of the
undesirable quality wine are found with a low citric acid content.

```{r echo=FALSE}
ggplot(aes(x=volatile.acidity, y=citric.acid, color = quality), data=rwine) + 
  labs(x = 'Volatile Acidity', y = 'Citric Acid',
       title = 'Volatile Acidity and Citric Acid by Quality') +
  scale_color_brewer(palette="Dark2", name = "Quality") +
  geom_point()
```

This graph shows us that a higher volatile acidity with low citric acid will
positively impact the quality of a wine. Wine that have high citric acidity and
a medium to low volatile acidity will score lower, which is very interesting
to see.

```{r echo=FALSE}
ggplot(aes(x=fixed.acidity, y=citric.acid, color = quality), data=rwine) + 
  labs(x = 'Fixed Acidity', y = 'Citric Acid',
       title = 'Fixed Acidity and Citric Acid by Quality') +
  scale_color_brewer(palette="PuBu", name = "Quality") +
  geom_point()
```

There is a positive relationship between the fixed acidity, citric acid and 
quality that is clearly demonstrated by this graph.


```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=fixed.acidity, y=citric.acid), data=rwine) + 
  labs(x = 'Fixed Acidity', y = 'Citric Acid',
       title = 'Fixed Acidity and Citric Acid by Quality') +
  facet_wrap(~rating) + 
  geom_smooth() +
  scale_color_brewer(palette="PuBu", name = "Quality") +
  geom_point(aes(color = quality), position = 'jitter', alpha = 0.5)
```

By further dividing them in their rating category, this relationship is 
demonstrated even further. Especially in the amazing rating, where both a high
citric acid and fixed acidity will lead to a much higher quality.

```{r echo=FALSE}
ggplot(aes(x=fixed.acidity, y=density, color = quality), data=rwine) + 
  labs(x = 'Fixed Acidity', y = 'Density',
       title = 'Fixed Acidity and Density by Quality') +
  scale_color_brewer(palette="Set1", name = "Quality") +
  geom_point()
```

This graph shows us that the denser the wine, the quality is lower. At the same 
time, it can be seen that the density and fixed acidity are once again clearly
correlated. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=fixed.acidity, y=density), data=rwine) + 
  labs(x = 'Fixed Acidity', y = 'Density',
       title = 'Fixed Acidity and Density by Rating') +
  scale_color_brewer(palette="Set1", name = "Rating") +
  geom_smooth(aes(color = rating))+
  geom_point(aes(color = rating), position = 'jitter', alpha = 0.2) 
```

From this graph, it can be clearly seen that the amazing category, which 
contains higher rating, tend to have a lesser density than the other category.

```{r echo=FALSE, message=FALSE, warning=FALSE,}
ggplot(aes(x=chlorides, y=sulphates, color = quality), data=rwine) + 
  labs(x = 'Chlorides', y = 'Sulphates',
       title = 'Chlorides and Sulphates by Quality') +
  scale_color_brewer(palette="Set1", name = "Quality") +
  geom_point(alpha = 0.5)
```

Zooming in on this graph and separating them by rating gives us this:

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=chlorides, y=sulphates, color = quality), data=rwine) + 
  labs(x = 'Chlorides', y = 'Sulphates',
       title = 'Chlorides and Sulphates by Quality') +
  scale_color_brewer(palette="Set1", name = "Quality") +
  xlim(0,0.2) +
  facet_wrap(~rating) + 
  geom_point(alpha = 0.5)
```

With this, it can be concluded that most of the wines rated the same has
first of all a very similar chlorides and sulphates level. Also that both
variables are positively correlated with quality. 

Last but not the least, taking a look at a graph of the highest correlated
variable, alcohol, with one that is the lowest, pH.

```{r echo=FALSE}
ggplot(aes(x=alcohol, y=pH, color = quality), data=rwine) + 
  labs(x = 'Alcohol', y = 'pH',
       title = 'Alcohol and pH by Quality') +
  scale_color_brewer(palette="Paired", name = "Quality") +
  geom_point()
```

As seen previously, pH does not have an impact on the quality of the wine. Even
if it is directly related to fixed acidity, which is related to volatile
acidity. This is can be told by the spread of different quality throughout the
different pHs. The only variable in play here is the alcohol content.



# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

In order to better analyze and understand the relationships between the 
variables and what combination contributes to a higher perceived quality, 
the bivariates analysis have been fitted with either their quality and/or 
rating.

- The alcohol and the acidic measurements (volatile acidity and citric acid) all
fit with what is expected, moreover, volatile acidity does not improve quality
passed a certain point, and it is the same with citric acid.
- Citric acid will provide the best wine at a mid-level, and will have an 
adverse effect if too high.
- Chlorides and sulphates are both positively correlated, and they both 
improve the quality of a wine.
- Higher quality wine appears less dense than their lower quality counterpart.


### Were there any interesting or surprising interactions between features?

- Although pH is directly affeced by the acidity of the wine, it does not have
a direct impact on the quality, and this across  different alcohol level, which
is the biggest correlator.
- Citric acid impact quality score more than the alcohol level, as with low 
citric acid,  there are a lot of lower tiered wine. But a high alcohol content
is always preferable.
- The chlorides and sulphates level of the wine  go hand in hand with their 
quality, and a lot of the wine will have a similar level across its rating.

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

N/A

------

# Final Plots and Summary

### Plot One

```{r echo=FALSE, Quality_v_Alcohol}
ggplot(aes(y=alcohol, x=quality), data=rwine) + 
  geom_jitter(alpha = 1/10) +
  geom_boxplot(alpha = 1/10, color = 'blue') +
  stat_summary(fun.y = 'mean', geom = 'point', color = 'red') + 
  labs(x = 'Quality', y = 'Alcohol Content (%)', title = 'Quality vs Alcohol')
```

### Description One

Alcohol, being the variable that is the highest correlated to quality, plays 
an important role in quality. A higher alcohol content will lead to a higher
quality in wine.

### Plot Two
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=chlorides, y=sulphates, color = quality), data=rwine) + 
  labs(x = 'Chlorides', y = 'Sulphates',
       title = 'Chlorides and Sulphates by Quality') +
  scale_color_brewer(palette="Set1", name = "Quality") +
  #xlim(0,0.2) +
  facet_wrap(~rating) + 
  geom_point(alpha = 0.5)
```

### Description Two
This graph is interesting as it demonstrates that no matter the quality of the
wine, the wine have a grouping of sulphates and chlorides. There are also quite
a few outliers, meaning that some wine can have alot of chlorides or  sulphates,
or both. But in that case, the wine tend to me of comparatively lesser quality
than those who are medium to low in each case.

### Plot Three
```{r echo=FALSE, message=FALSE, warning=FALSE, Citric_Acid_v_Alcohol_Rating}
ggplot(aes(y=alcohol, x=citric.acid), data=rwine) + 
  labs(y = 'Alcohol', x = 'Citric Acid',
       title = 'Citric Acid And Alcohol by Rating') +
  scale_color_brewer(palette="Set1", name = "Rating") +
  geom_smooth(aes(color = rating))+
  geom_point(aes(color = rating), position = 'jitter', alpha = 0.5) 
```

### Description Three

Firstly, it is undenaible that alcohol plays a huge role in the quality of the
wine. As a good wine will almost always have a high alcohol level. At the same 
time, although the trendlines shows otherwise, it can be seen from the points 
that the citric acid level also plays a big role. As most of the undersirable 
wines tends to a very low level of citric acid. From the above graph, it can be 
seen that the expert tend to seek a range in the citric acid, a too low or too
high concentration are not recommeneded. 
------

# Reflection

This is a dataset of 1,599 obervations of wine with 12 variables, which are 11 
input variables describing the different chemical properties of the wine and 1
output varible which is the quality of the wine. The quality is provided by wine
experts rating them from 0 (worst) to 10 (best).

The first step was to load the csv file and take a quick look to see its format
and have a feeling of the data and see where to start.

The first step was to do an univariate analysis to better understand each of the
variables. It can be seen that most of the quality are between 5 and 6. For the 
lack of data for the other quality, the rating variable has been created. Most 
of the variable follow a normal distribution, which indicates that wine of
similar quality will have a similar chemical components.

In further analysis, it has been shown that alcohol, volatile acidity, sulphates
and citric acid all play a bigger role in the quality of the wine, with alcohol
having the highest correlation. In brief, it can be concluded  that the 
likelyhood that a wine is considered good will have a higher alcohol content, 
lower volatile acidity, higher citric acid and higher sulphates. In this also 
then that it was discovered that despite pH being highly correlated to volatile
acidity, does not play a role in determining the quality of a wine.

The relationship between each of the variables has also be studied, where, as 
expected, volatile acidity and fixed acidity are all correlated 
with the pH level. And another obvious negative correlation is between density
and alcohol. More interestingly so, chlorides and sulphates are  also 
correlated, but with the majority of them being clustered together, with a few
outlier. This relationship leads to believe that most wine will seek the same
level in those 2 chemical components, with the outliers being a hit or miss.

The relationship between sulphates and chlorides have been studied further in
the multivariate analysis, where it can be seen that although they do have a 
direct impact on the quality, the outliers, in other words the wine with a high
amount of either, are often considered less desirable. Following the density 
analysis, it can be seen that better wine are less dense, which is to be 
expected given that alcohol have a smaller density than water. The relationship
between the volatile acidity and citric acid with quality definetively proved
that a good wine will have less volatile acidity and higher citric acid. Once
again, the  pH with the biggest correlator, alcohol, revealed that no matter the
pH, a higher alcohol will lead to a higher quality of wine.

Taking a look back, there were mny limitations to this analysis. First of all,
most of the wine's quality are 5 or 6. It would be beneficial to have a more
widespread quality ratings, to determine the relationships with higher
confidance. It would also be interesting to add where the wine is coming from,
the kind of grape being used, and also their price tag. As some people say, a
great wine does not need to be expensive, it would be interesting to verify
that claim. 